Niche complementarity among pollinators increases community-level plant reproductive success

### 2020-02-27 Network_structure_and_function.Rmd
changes for ecol lett

Declines in pollinator diversity and abundance have been reported across different regions, with implications for the reproductive success of plant species. However, research has focused primarily on pairwise plant-pollinator interactions, largely overlooking community-level dynamics. Here, we present one of the first efforts linking pollinator visitation to plant reproduction from a community-wide perspective using a well-replicated dataset encompassing 16 well-resolved plant-pollinator networks and data on reproductive success for 19 plant species from Mediterranean shrub ecosystems. We find that models including simple visitation metrics are sufficient to explain the variability in reproductive success observed. However, insights into the mechanisms through which differences in pollinator diversity translate into changes in reproductive success require additional information on network structure. Specifically, we find a positive effect of increasing niche complementarity between pollinators on plant reproductive success. This shows that maintaining communities with a diversity of species but also of functions is paramount to preserving natural ecosystems.

@@

Pollinators provide key services to plants by facilitating pollen flow between individuals. The recent declining trends found for some pollinator species in some regions of the planet ([@Potts2010; @I.2019]) have led many researchers to focus on the functional impacts of these changes in pollinator diversity, with a major focus being placed on the consequences for plant reproductive success [@Biesmeijer2006].

  Many research efforts have targeted the reproductive success of individual plant species [@Albrecht2012; @doi:10.1111/1365-2745.13055], and used relatively simple visitation metrics (e.g., the number of pollinator species that visit a plant or the number of visits they perform) to explain the differences observed across different plant individuals. Contrastingly, community-level analyses remain scarce [@Bennett2018]. Yet plants and pollinators do not interact in isolation, but rather are embedded within larger networks of interactions encompassing other plant and pollinator species. We are thus missing an important part of the picture, which includes the direct interactions between the whole ensemble of plants and pollinators, but also the indirect interactions between species within one guild (e.g., plants) through their shared resources [@Carvalheiro2014; @Lazaro2014; @Mayfield2017; @Pauw2013; @Johnson2019]. Understanding how changes in pollinator diversity and interaction structure affect whole community functioning is thus a major challenge that requires attention.

  The few pollination studies that have analysed the effects of changing pollinator diversity for reproductive success at the community level have done so using mainly experimental setups. As an example, a study that experimentally recreated a plant community with 9 plant species and differing levels of pollinator diversity across different enclosures, found that not only pollinator species diversity had an effect for average reproductive success, but that plant-pollinator interaction structure also had an important effect [@Frund2013]. In particular, these authors found that niche complementarity between pollinators, in terms of plant species and temperature coverage (a measure of the overlap in the use of plant resources and optimum temperature activity) had a positive effect for average seed set at the community level [@Frund2013]. This provides added information on the effects of biodiversity for ecosystem functioning, suggesting that not only the diversity of species present, but also the diversity of roles and ways in which a community is structured, are determinant factors.

  Indeed, theoretical research has long suggested that the structure of multitrophic communities has an effect for ecosystem functioning (reviewed in [@Thompson2012]). This line of research, rooted in niche theory and revamped by food-web studies [@Godoy2018; @Macarthur1967; @May1972a; @Tilman1982], has greatly advanced theory, but these ideas have not yet been tested using empirical data (but see [@Poisot2013]). Specifically, a major knowledge gap resides in understanding which aspects of structure determine which aspects of function [@Thompson2012]. This is because although a network perspective has promised to encapsulate complex ecological mechanisms occurring at the community level – such as indirect interactions [@Abrams1998; @Holt1977] or niche overlap [@Woodward2002]- less attention has been given to the ways in which these mechanisms relate to observed ecosystem processes [@Bluthgen2010a]. In contrast, we are now at a point in which there is considerable understanding on the attributes characterizing mutualistic interaction networks [@Bascompte2007a]. Especially, in the case of pollination, we have ample knowledge on the attributes that shape these mutualistic interactions at the community level. Amongst them is the prevalence of nested structures, i.e., arrangements where specialist species interact with a subset of the species that generalists interact with [@Bascompte2003] and which is thought to promote species diversity [@Bastolla2009]; or the relatively high extent of complementary specialization at the community scale, which may be directly related to key ecosystem functions [@Bluthgen2011]. However, the mechanisms by which these attributes affect plant reproduction remain to be understood [@Winfree2013]. The time is thus ripe to use the existing knowledge around plant-pollinator network structures to explore the relationship between network structure and ecosystem functioning empirically, with special emphasis being placed on the underlying ecological mechanisms that drive these relationships.

  Here, we present one of the first efforts linking pollinator visitation and plant reproductive success at the community level using empirical data on plant-pollinator interaction networks and plant reproductive success. To this end, we use a well-replicated dataset encompassing 16 well-resolved plant-pollinator interaction networks coupled with data on the reproductive success of 19 plant species recorded in Mediterranean shrub ecosystems. Our study focuses on understanding whether adding information on selected interaction network structure indices to previously used simple visitation metrics (e.g., the number and diversity of pollinator species visiting a plant species) aids in better explaining the differences observed in community-wide reproductive success. In doing so, we conducted our analyses focusing on reproductive success at two different levels: (i) at the species level by considering the effect of the position of a focal species within the larger network and its impact on its individual reproductive success, and (ii) at the site level, by evaluating how attributes that describe the whole site might affect average values of reproductive success. Specifically, our study focuses on how the interplay between niche complementarity and redundancy determines reproductive success. Plant reproductive success requires of the delivery of conspecific pollen and thus of a certain degree of niche complementarity [@Bluthgen2011]. Yet, greater values of nestedness, which imply redundancy in species functions, are thought to promote species diversity [@Bastolla2009] and stability [@Thebault2010a] within plant-pollinator networks. At present, we do not know how either of these network characteristics affects the functions performed by pollinators. Finally, in addition to average values, we also evaluate whether network structure helps explain differences in equity in reproductive success across species within a community, as a measure of evenness in the pollination service delivered.

@@

library(BeeFunData)#this package holds all the necessary data
library(plyr)
library(dplyr)
library(ggplot2)
library(grid)
library(iNEXT)
require(Hmisc)
library(knitr)
library(lme4)
library(lmerTest) #to calculate coeff and p values
library(car) #to calculate VIF
library(bbmle)  #to select best model based on AIC
library(ggplot2)
library(DHARMa) #to check for problems in residuals
library(reshape2)
library(spaa)
library(cowplot)
#devtools::install_github("hohenstein/remef")
library(remef)
library(glmmTMB)
library(vegan)
library(piecewiseSEM)
library(MuMIn)
library(jtools)

@@

d<-all_interactions
f<-fruitset
s<-seedset

@@

#calculate sorensen index to look at beta-diversity across sites for plants
library(betapart)
d.sub<-subset(d, d$Out=="transect") #keep only plants surveyed in transects, remove focals
d.beta<-table(d.sub$Site_ID, d.sub$Plant_gen_sp)
d.beta[d.beta>1]<-1 
d.betadiv<-mean(betadiver(d.beta, method = 11)) 
#summary(betadiver(d.beta, method = 11)) #max value 0.66.
```

@@

Our study was conducted in SW Spain within the area of influence of Doñana National Park, i.e., within the limits of the Natural Space of Doñana as defined by the local government (Junta de Andalucía, Fig. 1). All sites were located within similar elevations (ranging from 50 to 150 m a.s.l.), similar habitat and soil types, and presented similar plant composition (plant mean Sørensen beta-diversity among sites = `r round(d.betadiv, 2)`), reducing potential confounding factors. Here, we surveyed 16 Mediterranean woodland patches with an average distance of 7 km between them (min= 3 km, max= 46.5 km). Each site was surveyed 7 times during the flowering season of 2015 (from February to May) following a 100-m x 2 m transect for 30 mins. Along each transect, we identified all plant species and recorded all the floral visitors that landed on their flowers during each 30-min period. Only floral visitors (from now on referred to as pollinators) that could not be identified in the field were captured, stored and identified in the laboratory by FPM and experts in the different taxonomic groups (see acknowledgements). In addition, at each round we conducted 3 minutes of focal observations recording all floral visitors observed on 3 plant individuals per species belonging to the 19 most common (based on previous surveys) plant species across the study area (mean $\pm$ SD: `r round(mean(foc3$tot.species), digits=2)` $\pm$ `r round(sd(foc3$tot.species), digits=2)` species per site). Furthermore, we included some interactions between plant and pollinator individuals that were not observed during the sampling but that were opportunistically recorded immediately before or after the sampling periods, as some of these interactions are difficult to document and might be important to define network structure [@Jordano2016]. These opportunistic interactions represented `r round(table(d$Out)[2]/(table(d$Out)[1]+table(d$Out)[2]+table(d$Out)[3])*100, digits=2)` of all interactions recorded. All surveys were done under similar weather conditions, avoiding windy or rainy days. Surveys were done during mornings and afternoons with the sampling order being established randomly.

@@

s.fruitnum<-ddply(s, c("Site_ID", "Plant.sp"), summarise, 
                   tot.fruits = length(Plant.ID)) #number of fruits per individual
```

@@

Within each site, we marked between 3 and `r max(f.inds$tot.inds)` individuals (mean $\pm$ SD: `r round(mean(f.inds$tot.inds), digits=2)` $\pm$ `r round(sd(f.inds$tot.inds), digits=2)`, Table S2) belonging to `r min(tableS1$tot.sps)` to `r max(tableS1$tot.sps)` plant species (mean $\pm$ SD:`r round(mean(tableS1$tot.sps), digits=2)` $\pm$ `r round(sd(tableS1$tot.sps), digits=2)`, Table S1), depending on the availability and presence of flowers during the sampling events. For each individual, at the end of the season, we 
recorded fruit set (i.e. the proportion of flowers that set fruit), the average number of seeds per fruit and the average fruit and seed weight per fruit (`r min(s.fruitnum$tot.fruits)`-`r max(s.fruitnum$tot.fruits)` fruits subsampled, mean $\pm$ SD: `r round(mean(s.fruitnum$tot.fruits), digits=2)` $\pm$ `r round(sd(s.fruitnum$tot.fruits), digits=2)`, Table S3). Our survey included a total of `r nlevels(f2$Plant.sp)` different plant species across our `r nlevels(f2$Site_ID)` sites. Plants species were selected based on their availability, with sampling being focused on the most abundant plant species. The values at the species level were then averaged per site to calculate unique reproductive success measures at the site level. All plant species depend on pollinators to maximize their reproduction (Table S4).

@@

In order to evaluate the completeness of our sampling of the pollinator and plant community as well as that of their interactions, we estimated the asymptotic number of species present [@Chao2009], a non-parametric estimator of species richness for abundance data. This estimator includes non-detected species and allowed us to calculate the proportion detected with our original sampling data. We used Chao 1 asymptotic species richness estimators [@Chao2009] and estimated the richness of pollinators, plants and plant–pollinator links accumulated as sampling effort increased up to 100% sampling coverage using package iNEXT [@Hsieh2016] within the R environment [@RDevelopmentCoreTeam2011]. We then extracted the values covered by our sampling.

  In order to analyse how differences in network structure might affect plant reproductive success, we constructed plant-pollinator interaction networks by pooling the data for the 7 rounds of sampling. We thus obtained one interaction network per site, representing the number of individuals of different pollinator species recorded visiting each different plant species. For each network, we then proceeded to extract a series of relevant network metrics at the species and site levels.

  In addition, we checked for potential spatial autocorrelation in our data by means of Mantel correlograms. Autocorrelation values were low for all variables included in our analyses (Figure S1) and hence we treat each site as independent in our analysis.

@@

At the species level, we focused on attributes defining the position of a focal plant species within the larger community. As such, we considered two metrics providing complementary non-redundant information: (i) average niche overlap in terms of pollinators between a focal plant species and each of the other plant species in the community, which estimates the potential indirect interactions between different plant species through shared resources (in this case pollinators), and (ii) centrality, which depicts the importance of the role played by a plant species within the larger community (as resource for a large number of pollinator species) and its contribution to network cohesiveness. 
 Niche overlap was calculated as the average overlap in pollinator species visiting a focal plant and each of the other plants in the community using the Morisita overlap index, a measure of similarity between two sets of data [@Zhang2016]. As a measure of centrality, we used weighted closeness centrality, which represents the number of shortest paths going through a focal plant based on a unipartite projection of the bipartite plant-pollinator network using a weighted representation of the network [@Dormann2009]. With this method, links between plant species represent shared pollinator species.

@@

At the site level, we followed the same logic as the one presented at the species level. Thus, we also calculated two network metrics providing complementary non-redundant information. In this case we focused on (i) nestedness and (ii) pollinator niche complementarity.

  Nestedness is the property by which specialists interact with a subset of the species that generalists interact with [@Bascompte2003]. Although there is an ongoing debate in the literature, some studies have found that nested networks are more stable and resilient to perturbations because nestedness promotes a greater diversity by minimizing competition among species in a community [@Bastolla2009]. However, many network attributes vary with network size and complexity [@Bluthgen2006]. In the case of nestedness, we know it can be affected by network size and connectance [@Song2017a]. An approach that is often used to correct for this, is to use null models and to compare null-model corrected nestedness values across different networks. However, this approach has been recently shown to present the same issues, as z-scores also change with network size and connectance [@Song2017a]. We thus followed the advice provided by Song et al. (2017) by using a normalized value of the widely-used nestedness metric NODF [@Almeida-Neto2011], $NODF_c$. This normalized value is calculated as $NODF_c = NODF_n/(C * log(S))$, where C is connectance and S is network size. $NODF_n$ is calculated as $NODF/max(NODF)$, which is independent of network size and thus comparable across different networks [@Song2017a]. To calculate max(NODF) we used a recently corrected version of the algorithm [@Simmons2019a] in all but three sites, where the condition that the number of links>number of species was not met and thus precluded us from using this new version. Results did not change qualitatively when using the uncorrected version of the algorithm for all sites.

  Niche complementarity metrics are important because plant reproductive success depends on the delivery of conspecific pollen and thus of a certain level of specialization or niche divergence (reviewed in [@Brosi2016]). To calculate niche complementarity, we used a community-level measure defined as the total branch length of a functional dendrogram based on qualitative differences measured using a Euclidean distance in visitor assemblages between plants [@Devoto2012; @Petchey2007a]. All network metrics were calculated using package bipartite [@Dormann2009].

  All of these metrics were calculated using all the data as well as for the subset of the data excluding interactions observed outside of sampling periods. Differences between results are minimal for both and thus we will only present results for the analysis using the full dataset (see Table S12A-H for results removing observations out of transect).

@@

In order to evaluate whether adding information on network structure improves our ability to explain differences in reproductive success - both at the species and the site level - we used generalized linear (GLMs) and generalized linear mixed models (GLMMs) respectively. In both cases (species and site-level models) we fit two types of models: (i) model 1, that only included simple visitation metrics and (ii) model 2 that additionally included information on network structure. These models are meant to be additive, so that the network metrics included are intended to complement rather than substitute the simple metrics traditionally used.

  At the species level, response variables included the fruit set for different individuals of each species analyzed using a binomial distribution and the average number of seeds per fruit, the average fruit and seed weight fitted using normal distributions. The number of seeds per fruit was centered and scaled (i.e., we subtracted column means and divided by standard deviation) to allow meaningful comparisons across species with contrasting life histories. As explanatory variables, model 1 included pollinator richness, and the total number of visits received by each plant species; while model 2 added the two network attributes calculated at the species level: average plant niche overlap and centrality. For both models, we included plant species nested within site and site as random effects to account for the non-independence of several individuals measured per species and site.

	At the site level, we upscaled our species-level analyses. As response variables we had the average reproductive success per site (i.e., average fruit set analyzed using a binomial distribution, average number of seeds per fruit and average fruit and seed weight using a normal distribution). We thus had a single value per site and no random effects are needed. In this case, model 1 included total pollinator richness and total pollinator abundance (i.e. number of visits received by all plants within the community) as explanatory variables. Model 2, in turn, added information on network structure by including nestedness and pollinator niche complementarity as explanatory variables.

  Average values of reproductive success at the site level can be driven to a large extent by a single plant species. Yet, what will determine the persistence of a diverse plant community, is the presence of some sort of “equity” or evenness in reproductive success across the whole community. We therefore calculated a measure of equity in reproductive success at the site level as the proportion of species with normalized (between 0 and 1) average fruit set values that were above the 50^th percentile. As any selected threshold is arbitrary, we repeated this using the 25^th and 75^th percentile thresholds [@Byrnes2014]. We then used the same framework as that used for species and site-level analyses and fit the same models 1 and 2 GLMs but using equity in reproductive success as the response variable and fitting a binomial distribution.

  In all cases, we used variance inflation factors to check for collinearity between explanatory variables. Additionally, we ran residual diagnostics to check if model assumptions were met. Then, we used the Akaike Information Criterion (AIC) to compare model performance and complexity. Whenever the difference between the AIC of both models was < 2 ($\Delta AIC<2$), we considered that both models were equally good [@Burnham2011]. All predictor variables were standardized prior to analysis. For every model we also calculate the R2 value using the approximation suggested for generalized mixed models when necessary [@Shinichi2017].

  Finally, we tested whether the importance of network structure in explaining differences in equity in reproductive success within communities increases with the number of plant species being considered. We expect that when only one plant species is considered, then the importance of network structure will be negligible, while we expect this importance to increase as more plant species are considered (up to a maximum number of 6 species which is the maximum we have measured in our study at a particular site).

  To test this, we ran a simple simulation in which the number of species considered increased at each step and for each step we re-calculated equity in reproductive success. Instead of drawing plant species randomly for each step, we tested all possible combinations for each plant number level and network, as the number of combinations is small (e.g. for n = 3 plant selected out of 6 there is only 20 possible combinations). Then, we tested if the relationship between equity in reproductive success and functional complementarity (given its importance in determining differences in reproductive success, see Results section) changes as a function of the number of plants considered within our simulated communities. To this end, for each level of species number considered, we randomly selected one of the generated equity values across each of the 16 communities and regressed these 16 values against our network level predictor and extracted the model slope estimates. We repeated this process 1,000 times and averaged all slope estimates. We expect that the more plants considered, the larger the resulting average estimates will be. Note that we only interpret the mean effects, as the variance among different plant number of species considered depends on the initial number of possible combinations.

@@

#check proportion represented by each order, hymenoptera, diptera, coleoptera...
 tr<-read.csv("traits_ainhoa.csv")
# #head(tr)
# 
 d$order<-tr$order[match(d$Pollinator_gen_sp, tr$Pollinator_gen_sp)]
 d.order<- aggregate(Frequency ~ order, d, FUN= "sum")
 d.order$prop<-d.order$Frequency/sum(d.order$Frequency)

@@

sites <- unique(d$Site_ID)

@@

AICtab(m1, m2)

>>>>

The number of studies analyzing the structure of interactions shaping natural communities has increased, particularly regarding pollination. However, research linking pollinator diversity to reproductive success has focused on pairwise plant-pollinator interactions, largely overlooking community-level dynamics. Here, we present one of the first single studies linking pollinator visitation to plant reproduction from a community-wide perspective using a well-replicated dataset encompassing 16 plant-pollinator networks and data on reproductive success for 19 plant species from Mediterranean shrub ecosystems. We find that statistical models including simple visitation metrics are sufficient to explain the variability observed. However, a mechanistic understanding of how pollinator diversity affects reproductive success require additional information on network structure. Specifically, we find positive effects of increasing complementarity in the plant species visited by different pollinators on plant reproductive success. This shows that maintaining communities with a diversity of species but also of functions is paramount to preserving plant diversity.

@@

Pollinators provide key services to plants by facilitating pollen flow between individuals. Records of declining trends for some pollinator species in some regions ([@Potts2010; @I.2019]) have led researchers to focus on the functional impacts of these changes in pollinator diversity, especially on the consequences for plant reproductive success [@Biesmeijer2006].

  Many research efforts have targeted the reproductive success of individual plant species [@Albrecht2012; @doi:10.1111/1365-2745.13055], and used relatively simple visitation metrics (e.g., the number of pollinator species that visit a plant or the number of visits they perform) to explain the differences observed. Contrastingly, community-level analyses remain scarce [@Bennett2018]. Yet plants and pollinators do not interact in isolation, but rather are embedded within larger networks of interactions encompassing other plant and pollinator species. We are thus missing an important part of the picture, which includes the direct interactions between the whole ensemble of plants and pollinators, but also the indirect interactions between species within one guild (e.g., plants) through their shared resources [@Carvalheiro2014; @Lazaro2014; @Mayfield2017; @Pauw2013; @Johnson2019]. Understanding how changes in pollinator diversity and interaction structure affect whole ecosystem functioning is thus a major challenge that requires attention.

  The few pollination studies that have analysed the effects of changing pollinator diversity for reproductive success at the community level have done so using mainly experimental setups. As an example, a study that experimentally recreated a plant community with 9 plant species and differing levels of pollinator diversity across different enclosures, found a positive effect of pollinator species diversity on seed set, but also an important effect of niche complementarity between pollinators, a measure of community structure [@Frund2013]. This provides added information on the effects of biodiversity for ecosystem functioning, suggesting that not only the diversity of species present, but also the diversity of roles and ways in which a community is structured, are determinant factors.

  Indeed, theoretical research has long suggested that the structure of multitrophic communities has an effect for ecosystem functioning (reviewed in [@Thompson2012]). This line of research, rooted in niche theory and revamped by food-web studies [@Godoy2018; @Macarthur1967; @May1972a; @Tilman1982], has greatly advanced theory, but the relationship between structure and function has seldom been tested using empirical data (but see [@Kaiser-Bunbury2017; @Poisot2013; @Saunders2019]). This line of research, rooted in niche theory and revamped by food-web studies  has greatly advanced theory, but these ideas have not yet been tested using empirical data  Specifically, a major knowledge gap resides in understanding which aspects of structure determine which aspects of function [@Thompson2012]. This is because although a network perspective has promised to encapsulate complex ecological mechanisms occurring at the community level – such as indirect interactions [@Abrams1998; @Holt1977] or niche overlap [@Woodward2002]- less attention has been given to the ways in which these mechanisms relate to observed ecosystem processes [@Bluthgen2010a]. In contrast, we are now at a point in which there is considerable understanding on the attributes characterizing mutualistic interaction networks [@Bascompte2007a]. Especially, in the case of pollination, we have ample knowledge on the attributes that shape these mutualistic interactions at the community level. Amongst them is the prevalence of nested structures, i.e., arrangements where specialist species interact with a subset of the species that generalists interact with [@Bascompte2003] and which is thought to promote species diversity [@Bastolla2009]. Further, plant-pollinator interaction networks seem to exhibit a relatively high extent of complementary specialization at the community scale, which may be directly related to key ecosystem functions [@Bluthgen2011]. However, the mechanisms by which these attributes affect plant reproduction remain to be understood [@Winfree2013]. The time is thus ripe to use the existing knowledge around plant-pollinator network structures to explore the relationship between network structure and ecosystem functioning empirically, with special emphasis on the underlying ecological mechanisms that drive these relationships.

  Here, we present one of the first efforts linking pollinator visitation and plant reproductive success at the community level using empirical data on plant-pollinator interaction networks and plant reproductive success. To this end, we use a well-replicated dataset encompassing plant-pollinator interaction networks collected at 16 sites coupled with data on the reproductive success of 19 plant species recorded in Mediterranean shrub ecosystems. Our study focuses on understanding whether adding information on selected interaction network structure indices to previously used simple visitation metrics (e.g., the number and diversity of pollinator species visiting a plant species) aids in better explaining the differences observed in community-wide reproductive success. In doing so, we conducted our analyses focusing on reproductive success at two different levels: (i) at the species level by considering the association between the position of a focal species within the larger network and its link to individual reproductive success, and (ii) at the site level, by evaluating how attributes that describe the whole site might affect average values of reproductive success for all species measured within one particular site. Specifically, our study focuses on how the interplay between the complementarity in plant species visited by different pollinators (i.e., the complementarity in the functions they perform) and the redundancy in this function relates to reproductive success. Plant reproductive success requires of the delivery of conspecific pollen and thus of a certain degree of niche complementarity [@Bluthgen2011]. Yet, greater values of redundancy in species functions (e.g., that provided by nested structures), are thought to promote species diversity [@Bastolla2009] and stability [@Thebault2010a] within plant-pollinator networks. At present, we do not know how either of these network characteristics affects the functions performed by pollinators.

@@

library(BeeFunData)#this package holds all the necessary data
library(plyr)
library(dplyr)
library(ggplot2)
library(grid)
library(iNEXT)
require(Hmisc)
library(knitr)
library(lme4)
library(lmerTest) #to calculate coeff and p values
library(car) #to calculate VIF
library(bbmle)  #to select best model based on AIC
library(ggplot2)
library(DHARMa) #to check for problems in residuals
library(reshape2)
library(spaa)
library(cowplot)
#devtools::install_github("hohenstein/remef")
library(remef)
library(glmmTMB)
library(vegan)
library(piecewiseSEM)
library(MuMIn)
library(jtools)
library(sjPlot)

@@

d<-all_interactions
dtrans<-subset(d, d$Out=="transect")  #keep only records within transects, eliminate focal and other records outside transects
f<-fruitset
s<-seedset

@@

#calculate sorensen index to look at beta-diversity across sites for plants
library(betapart)
dtrans<-subset(d, d$Out=="transect") #keep only plants surveyed in transects, remove focals
d.beta<-table(dtrans$Site_ID, dtrans$Plant_gen_sp)
d.beta[d.beta>1]<-1 
d.betadiv<-mean(betadiver(d.beta, method = 11)) 
#summary(betadiver(d.beta, method = 11)) #max value 0.66.
```

@@

Our study was conducted in SW Spain within the area of influence of Doñana National Park, i.e., within the limits of the Natural Space of Doñana as defined by the local government (Junta de Andalucía, Fig. 1). All 16 sites were located within similar elevations (ranging from 50 to 150 m a.s.l.), similar habitat and soil types, and presented similar plant composition (plant mean Sørensen beta-diversity among sites = `r round(d.betadiv, 2)`), reducing potential confounding factors. Here, we surveyed 16 Mediterranean woodland patches with an average distance of 7 km between them (min= 3 km, max= 46.5 km). Each site was surveyed every two weeks for a total of 7 times during the flowering season of 2015 (from February to May) following a 100-m x 2 m transect for 30 mins. Along each transect, we identified all plant species and recorded all the floral visitors that landed on their flowers during each 30-min period. Only floral visitors (from now on referred to as pollinators) that could not be identified in the field were captured, stored and identified in the laboratory by FPM and another expert entomologist (see acknowledgements). All surveys were done under similar weather conditions, avoiding windy or rainy days, during mornings and afternoons with the sampling order being established randomly.

@@

s.fruitnum<-ddply(s, c("Site_ID", "Plant.sp"), summarise, 
                   tot.fruits = length(Plant.ID)) #number of fruits per individual

@@

Within each site, we marked between 3 and `r max(f.inds$tot.inds)` individuals (mean $\pm$ SD: `r round(mean(f.inds$tot.inds), digits=2)` $\pm$ `r round(sd(f.inds$tot.inds), digits=2)`, Table S2) belonging to `r min(tableS1$tot.sps)` to `r max(tableS1$tot.sps)` plant species (mean $\pm$ SD:`r round(mean(tableS1$tot.sps), digits=2)` $\pm$ `r round(sd(tableS1$tot.sps), digits=2)`, Table S2). For each individual, at the end of the season, we recorded fruit set (i.e. the proportion of flowers that set fruit), the average number of seeds per fruit and the average fruit and seed weight per fruit (`r min(s.fruitnum$tot.fruits)`-`r max(s.fruitnum$tot.fruits)` fruits subsampled, mean $\pm$ SD: `r round(mean(s.fruitnum$tot.fruits), digits=2)` $\pm$ `r round(sd(s.fruitnum$tot.fruits), digits=2)`, Table S3). These variables show strong correlations (Pearson correlation= 0.89), we believe they provide complementary information. Our survey included a total of `r nlevels(f2$Plant.sp)` different totally or partially self-incompatible plant species that depend on pollinators to maximize their reproduction (Table S4) across our `r nlevels(f2$Site_ID)` sites. All plant species were shrubs selected to be common and widespread species, and individuals were selected depending on the presence of flowers during the sampling events. The values of reproductive success obtained at the species level were also averaged per site to calculate unique reproductive success measures at the site level in addition to species-level values..

@@

In order to evaluate the completeness of our sampling, we estimated the asymptotic number of species of plants, pollinators and interactions present [@Chao2009], a non-parametric estimator of species richness for abundance data. This estimator includes non-detected species and allowed us to calculate the proportion detected with our original sampling data. We used Chao 1 asymptotic species richness estimators [@Chao2009] and estimated the richness of pollinators, plants and plant–pollinator links accumulated as sampling effort increased up to 100% sampling coverage using package iNEXT [@Hsieh2016] within the R environment [@RDevelopmentCoreTeam2011]. We then extracted the values covered by our sampling.

  In order to analyse how differences in network structure between communities might affect plant reproductive success, we constructed quantitative plant-pollinator interaction networks by pooling the data for the 7 rounds of sampling. We thus obtained one interaction network per site, representing the number of individuals of different pollinator species recorded visiting each different plant species. For each network, we then extracted a series of relevant network metrics at the species and site levels.

  In addition, we checked for potential spatial autocorrelation in our data by means of Mantel correlograms. Autocorrelation values were non-significant for all variables included in our analyses, except in the case of pollinator richness where we have a significant effect at small spatial scales (Fig. S1), and hence we treat each site as independent in our analysis.

@@

At the species level, we focused on attributes defining the position of a focal plant species within the larger community. As such, we considered two metrics providing complementary non-redundant information: (i) average niche overlap in terms of pollinators between a focal plant species and each of the other plant species in the community, which estimates the potential indirect interactions between different plant species through shared resources (in this case pollinators) and the potential for increased heterospecific pollen [@Arceo-Gomez2019], and (ii) centrality, which depicts the importance of the role played by a plant species within the larger community (as resource for a large number of pollinator species) and its contribution to network cohesiveness. 
 Niche overlap was calculated as the average overlap in pollinator species visiting a focal plant and each of the other plants in the community using the Morisita overlap index, a measure of similarity between two sets of data [@Zhang2016]. As a measure of centrality, we used weighted closeness centrality, which represents the number of shortest paths going through a focal plant based on a unipartite projection of the bipartite plant-pollinator network using a weighted representation of the network [@Dormann2009]. With this method, links between plant species represent shared pollinator species.

@@

At the site level, we followed the same logic as the one presented at the species level. Thus, we also calculated two network metrics providing complementary non-redundant information. In this case we focused on a metric that measures the redundancy in the plants visited by different pollinators, nestedness, and another one, not correlated with the previous one, measuring the complementarity in the plant species visited by the different pollinator species, pollinator niche complementarity

  Nestedness is the property by which specialists interact with a subset of the species that generalists interact with [@Bascompte2003]. Although there is an ongoing debate in the literature (e.g., [@James2012]), some studies have found that nested networks are more stable and resilient to perturbations because nestedness promotes a greater diversity by minimizing competition among species in a community [@Bastolla2009]. However, many network attributes vary with network size and complexity [@Bluthgen2006]. In the case of nestedness, we know it can be affected by network size and connectance [@Song2017a]. An approach that is often used to correct for this, is to use null models and to compare null-model corrected nestedness values across different networks. However, this approach has been shown to present the same issues, as z-scores also change with network size and connectance [@Song2017a]. We thus followed the advice provided by Song et al. (2017) by using a normalized value of the widely-used nestedness metric NODF [@Almeida-Neto2011], $NODF_c$. This normalized value is calculated as $NODF_c = NODF_n/(C * log(S))$, where C is connectance and S is network size, calculated as $S = \sqrt(ncol(web)*nrow(web))$. $NODF_n$ is calculated as $NODF/max(NODF)$, which is independent of network size and thus comparable across different networks [@Song2017a]. To calculate max(NODF) we used a corrected version of the algorithm [@Simmons2019a] whenever possible. Results did not change qualitatively when using the uncorrected version of the algorithm for all sites.

 To calculate niche complementarity, we used a community-level measure defined as the total branch length of a dendrogram based on qualitative differences in visitor assemblages between plants [@Devoto2012; @Petchey2007a]. All network metrics were calculated using package bipartite [@Dormann2009].

@@

To evaluate whether adding information on network structure improves our ability to explain differences in reproductive success - both at the species and the site level - we used generalized linear (GLMs) and generalized linear mixed models (GLMMs) respectively. In both cases (species and site-level models) we fit three types of models: (i) model 0, a null model with no explanatory variables,(ii) model 1, that only included simple visitation metrics and (iii) model 2 that additionally included information on network structure. These models are meant to be additive, so that the network metrics included are intended to complement rather than substitute the simple metrics traditionally used.

At the species level, response variables included fruit set analyzed using a binomial distribution and the average number of seeds per fruit, the average fruit and seed weight fitted using normal distributions. The number of seeds per fruit was centered and scaled (i.e., we subtracted column means and divided by standard deviation) to allow meaningful comparisons across species with contrasting life histories. As explanatory variables, model 1 included pollinator richness, and the relative number of visits received by each plant species, which is calculated as the total number of visits received by a plant species divided by the average number of flowers present in that species. In turn, model 2 added the two network attributes calculated at the species level: average plant niche overlap and centrality. For both models, we included plant species identity nested within site and site as random effects to account for the fact that several individuals of the same plant species were measured at each site.

	At the site level, response variables were the average reproductive success of all plants surveyed within a site (i.e., average fruit set analyzed using a binomial distribution, average number of seeds per fruit and average fruit and seed weight using a normal distribution). We thus had a single value per site and no random effects are needed. In this case, model 1 included total pollinator richness and total pollinator abundance (i.e. number of visits received by all plants within the community) as explanatory variables. Model 2, in turn, added information on network structure by including nestedness and pollinator niche complementarity.

  Average values of reproductive success at the site level can be driven to a large extent by a single plant species. Yet, what will determine the persistence of a diverse plant community, is the presence of some sort of “equity” or evenness in reproductive success across the whole community. We therefore calculated the proportion of species with normalized (between 0 and 1) average fruit set values that were above the 50^th percentile as a measure of equity. As any selected threshold is arbitrary, we repeated this using the 25^th and 75^th percentile thresholds [@Byrnes2014]. We then used the same framework as that used for species and site-level analyses and fit the same models 0, 1 and 2 GLMs using equity in reproductive success as response variable and fitting a binomial distribution.

  In all cases, we used variance inflation factors to check for collinearity between explanatory variables. Additionally, we ran residual diagnostics to check if model assumptions were met and used the Akaike Information Criterion (AIC) to compare model performance and complexity. Whenever the difference between the AIC of the models was < 2 ($\Delta AIC<2$), we considered that all models were equally good [@Burnham2011]. In the case of mixed models, for comparison, models were fitted by maximum likelihood and then the best model was refitted using restricted maximum likelihood. All predictor variables were standardized prior to analysis. For every model we also calculate the R2 value using the approximation suggested for GLMMs when necessary [@Shinichi2017].

  Finally, we tested whether the importance of network structure in explaining differences in equity in reproductive success increases with the number of plant species being considered. We expect that when only one plant species is considered the importance of network structure will be negligible, while we expect it to increase as more plant species are considered (up to a maximum number of 6 species which is the maximum we have measured in our study at a particular site).

  To test this, we ran a simple simulation in which the number of species considered increased at each step and for each step we re-calculated equity in reproductive success. Instead of drawing plant species randomly for each step, we tested all possible combinations for each plant number level and network, as the number of combinations is small (e.g. for n = 3 plant selected out of 6 there is only 20 possible combinations). Then, we tested if the relationship between equity in reproductive success and niche complementarity (given its importance in determining differences in reproductive success, see Results section) changes as a function of the number of plants considered within our simulated communities. To this end, for each level of species number considered, we randomly selected one of the generated equity values across each of the 16 communities and regressed these 16 values against our network level predictor and extracted the model slope estimates. We repeated this process 1,000 times and averaged all slope estimates. We expect that the more plants considered, the larger the resulting average estimates will be. Note that we only interpret the mean effects, as the variance among different plant number of species considered depends on the initial number of possible combinations.

@@

#check proportion represented by each order, hymenoptera, diptera, coleoptera...
 tr<-read.csv("traits_ainhoa.csv")
# #head(tr)
# 
 dtrans$order<-tr$order[match(dtrans$Pollinator_gen_sp, tr$Pollinator_gen_sp)]
 d.order<- aggregate(Frequency ~ order, dtrans, FUN= "sum")
 d.order$prop<-d.order$Frequency/sum(d.order$Frequency)

@@

sites <- unique(dtrans$Site_ID)

@@

AICtab(m0,m1, m2)
